---
title: "Palmer Penguins"
subtitle: "An example dashboard"
logo: "logo.png"
theme:
  - flatly
format: 
  dashboard:
    orientation: columns
---

```{r}
#| output: FALSE
#| echo: FALSE

# Load libraries 
library(shiny) 
library(palmerpenguins) 
library(tidyverse) 
library(DT) 
library(plotly)

```



# Dashboard

## Column

### Row


```{r}
#| component: valuebox

list(
  value = nrow(penguins),
  title = "Number of penguins observed",
  color = "blue",
  icon = "bar-chart"
)

```

### Row

```{r}
#| component: valuebox

list(
  value = length(unique(penguins$species)),
  title = "Number of species in sample",
  color = "green",
  icon = "pie-chart"
)
```


### Row

```{r}
#| component: valuebox

list(
  value = length(unique(penguins$island)),
  title = "Number of islands in sample",
  color = "red",
  icon = "pie-chart"
)
```



## Column {.tabset}

```{r}
#| title: Flipper Length

# make an interactive ggplot scatterplot of the flipper length by species
p <- ggplot(penguins, aes(x = flipper_length_mm, fill = species)) +
  geom_histogram(binwidth = 2, alpha = 0.64) +
  labs(title = "Flipper length by species",
       x = "Flipper length (mm)",
       y = "Count",
       fill = "Species") +
  scale_fill_manual(values = c("#1b9e77", "#d95f02", "#7570b3")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)
  )

# make the plot interactive
plotly::ggplotly(p)

```


```{r}
#| title: Body Mass

# Violin plot of body mass by species with semi-transparent colors (colors should be green, orange, and purple)

p2 <- ggplot(penguins, aes(x = species, y = body_mass_g, fill = species)) +
  geom_violin(alpha = 0.7) +
  labs(title = "Body mass by species",
       x = "Species",
       y = "Body mass (g)",
       fill = "Species") +
  scale_fill_manual(values = c("#1b9e77", "#d95f02", "#7570b3")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)
  )


# make the plot interactive
plotly::ggplotly(p2)

```


```{r}
#| title: Bill Length vs. Flipper Length

# Scatterplot of bill length vs flipper length colored by species (colors should be green, orange, and purple)

p3 <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point(alpha = 0.7) +
  labs(title = "Scatterplot of bill length vs flipper length",
       x = "Bill length (mm)",
       y = "Flipper length (mm)",
       color = "Species") +
  scale_color_manual(values = c("#1b9e77", "#d95f02", "#7570b3")) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)
  )

# make the plot interactive
plotly::ggplotly(p3)

```


# Data

## Column {width=20%}
You can view the data that was used to create this dashboard in the table. Click the button to download the data as a CSV file.

## Column {width=80%}

```{r}
datatable( palmerpenguins::penguins, 
           extensions = 'Buttons', 
           options = 
             list( 
               dom = 'Bfrtip', 
               buttons = list( 
                 list(extend = 'csv', text = 'Click here to download CSV') ) 
               ) 
           )

```


# About {width=70%}

The Palmer Penguins dataset is a popular dataset for data exploration and visualization examples. It contains data on three species of penguins collected from three islands in the Palmer Archipelago, Antarctica. 

Data was collected by [Dr. Kristen Gorman](https://www.uaf.edu/cfos/people/faculty/detail/kristen-gorman.php) and the Palmer Station, Antarctica LTER. 

The dataset was made available by the `palmerpenguins` R package,  developed by Allison Horst, Alison Hill, and Dr. Kristen Gorman.

For more information about the dataset and the package, you can visit the [package page](https://allisonhorst.github.io/palmerpenguins/). 



